### 2025-02-12 latex/regularization_comparison.tex
tikz

\begin{tikzpicture}
\begin{axis}[
    width=0.9\textwidth,
    height=6cm,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    xlabel style={font=\tiny},
    ylabel style={font=\tiny},
    tick label style={font=\tiny},
    title style={font=\small},
    legend style={font=\tiny, at={(0.02,0.98)}, anchor=north west},
    xlabel={Underlying Asset Price},
    ylabel={Payout},
    title={Regularization Methods Comparison}
]

\addplot[thick, black] table[x index=0,y index=1] {regularization_comparison.dat};
\addlegendentry{Target Payout}

\addplot[thick, dashed, red] table[x index=0,y index=2] {regularization_comparison.dat};
\addlegendentry{L2 Regularization Approximation}

\addplot[thick, dashdotted, blue] table[x index=0,y index=3] {regularization_comparison.dat};
\addlegendentry{L1 Regularization Approximation}

\end{axis}
\end{tikzpicture}