### 2025-02-19 latex/article.tex
alternative

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\onehalfspacing
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{amsmath, amssymb, amsthm}

% Define keywords command
\providecommand{\keywords}[1]{\textbf{Keywords:} #1}

% Remove page number from the first page
\usepackage{titling}
\usepackage{geometry}

% Adjust vertical spacing on title page
\pretitle{\vspace*{2cm}\begin{center}}
\posttitle{\end{center}\vspace{0.5cm}}
\preauthor{\begin{center}}
\postauthor{\end{center}\vspace{0.5cm}}
\predate{\begin{center}}
\postdate{\end{center}\vspace{1cm}}

\title{\rule{\textwidth}{4pt}\\[1em]
\textbf{From Distribution Forecasts to Vanilla Portfolios: \\[0.2cm]
Regularized Payout Approximation in Discrete Space}\\[1em]
\rule{\textwidth}{2pt}}

\author{\small \textbf{Vlad Kochetov} \\
    \small Faculty of Mechanics and Mathematics, \\
    \small Taras Shevchenko National University of Kyiv \\
    \small \texttt{vladkoch@knu.ua}
}
\date{\today}

\begin{document}

% Remove page number from the first page
\thispagestyle{empty}

\maketitle

\begin{abstract}
This paper explores the methodology for approximating complex payout profiles using a
combination of vanilla European options. The mathematical framework is presented that
allows for the decomposition of arbitrary payout structures into a series of call and
put options with varying strike prices. The approach is validated through numerical
examples and practical applications in financial engineering. The results demonstrate
the effectiveness of this method in replicating complex payouts.
\end{abstract}

\newpage

\keywords{
    \small{
        Vanilla Options, 
        Payout Profile, 
        Financial Engineering, 
        Option Pricing, 
        Approximation, 
        Numerical Methods,
        Quantitative Finance,
        Financial Derivatives,
        Computational Finance,
        Distribution Forecasts,
        Option Trading
    }
}

\section{Introduction}
The ability to replicate complex financial instruments using simpler, more liquid
derivatives is a cornerstone of modern financial engineering. Vanilla European 
options, due to their simplicity and widespread availability, serve as ideal 
building blocks for such approximations. This paper introduces an approach to 
approximating arbitrary payout profiles using a combination of these options.

\section{Literature Review}
The theoretical foundations of option portfolio optimization were established in the work of 
\citep{Carr01012001}, which demonstrated the theoretical possibility of replicating 
arbitrary payout functions through continuous strike spaces. While this approach 
provides elegant closed-form solutions in continuous market settings, several fundamental 
limitations affect practical implementation: the discrete availability of strike prices 
in real markets, liquidity constraints, non-negligible transaction costs scaling with 
portfolio complexity, and numerical stability challenges in finite difference implementations. 
The proposed method addresses these limitations through regularization techniques and 
discrete optimization methods tailored for real-world trading environments, while 
preserving the theoretical elegance of the original framework.

\section{Methodology}
\subsection{Model Assumptions}
The proposed approximation framework relies on several key assumptions:

\begin{enumerate}
    \item \textbf{Discrete Strike Space:} The model operates in a discrete strike price space 
    $\{K_i\}_{i=1}^n$, reflecting real market conditions where options are available only 
    at specific strikes.

    \item \textbf{Local Approximation Region:} The approximation focuses on a finite price 
    interval around the current spot price $[S_{\min}, S_{\max}]$, where $S_{\min}$ and $S_{\max}$ 
    are chosen to ensure sufficient coverage of the relevant price region.

    \item \textbf{European-Style Options:} The framework utilizes European-style vanilla 
    options exclusively, avoiding the complexity of early exercise features present in 
    American options.

    \item \textbf{Static Replication:} The model assumes a static replication approach, 
    where the portfolio weights remain constant until maturity. Dynamic rebalancing effects 
    are not considered in the basic framework.

    \item \textbf{Put-Call Parity Flexibility:} The implementation supports both direct 
    use of calls and puts as separate instruments, as well as expressing puts through 
    calls and spot positions using put-call parity. This approach provides flexibility 
    in portfolio construction while maintaining mathematical equivalence.
    This flexibility enables optimization of the portfolio structure based on market 
    conditions.
\end{enumerate}

The numerical implementation discretizes the approximation domain into a finite set of 
evaluation points. The number of discretization points is chosen to balance computational 
efficiency with approximation accuracy.

\subsection{Mathematical Framework}
Let \( V(S) \) represent the value of a target payout profile as a function of the 
underlying asset price \( S \). We construct an approximation using \( n \) call 
options and \( m \) put options with strike prices \( \{K_i\}_{i=1}^{n+m} \):

\begin{equation}
V(S) \approx \lambda S + 
    \sum_{i=1}^{n} \alpha_i C(S, K_i) + 
    \sum_{j=1}^{m} \beta_j P(S, K_j)
\end{equation}

where \( \lambda \) represents the spot position, \( \alpha_i \) and \( \beta_j \) 
are option weights. The optimal weights are found by solving the regularized least 
squares problem with L2 (Ridge) regularization:

\begin{equation}
\min_{\boldsymbol{\theta}} 
    \int_{S_{\min}}^{S_{\max}} 
        \left[ V(S) - \hat{V}(S;\boldsymbol{\theta}) \right]^2 dS + 
    \gamma \|\boldsymbol{\theta}\|_2^2
\end{equation}

where \( \boldsymbol{\theta} = (\lambda, \{\alpha_i\}, \{\beta_j\}) \) represents
the position weights, \( \gamma \) is the regularization parameter, and 
\( S_{\min}, S_{\max} \) define the approximation domain. The L2 norm promotes 
smoother weight distributions, providing stability to the solution.

Discretizing the integral and using matrix notation, we obtain:

\begin{equation}
\boldsymbol{\theta}^* = 
    \argmin_{\boldsymbol{\theta}} 
        \|\mathbf{A}\boldsymbol{\theta} - \mathbf{b}\|_2^2 + 
        \gamma \|\boldsymbol{\theta}\|_2^2
\end{equation}

where the design matrix \( \mathbf{A} \) contains option payoffs and spot positions
evaluated at discrete price points \( \{S_k\}_{k=1}^N \), and \( \mathbf{b} \) is
the vector of target payout values.

This quadratic optimization problem admits an analytical solution:
\begin{equation}
\boldsymbol{\theta}^* = 
    (\mathbf{A}^\top \mathbf{A} + \gamma \mathbf{I})^{-1} 
    \mathbf{A}^\top \mathbf{b}
\end{equation}

\subsection{Alternative Loss Functions}
While the standard L2 loss function provides analytical tractability, alternative 
formulations may better reflect trading objectives. Two notable generalizations are:

\begin{enumerate}
    \item \textbf{L1 Regularization (Lasso):} Replacing L2 regularization term with L1 norm:
    \begin{equation}
        L(\boldsymbol{\theta}) = 
            \int_{S_{\min}}^{S_{\max}} 
                \left[ V(S) - \hat{V}(S;\boldsymbol{\theta}) \right]^2 dS + 
            \gamma \|\boldsymbol{\theta}\|_1
    \end{equation}
    This formulation tends to produce sparse solutions by setting some weights exactly 
    to zero, which can be beneficial when portfolio simplicity is important consideration.

    \item \textbf{Weighted Error Function:} Incorporating target function magnitude 
    into the error term:
    \begin{equation}
        L(\boldsymbol{\theta}) = 
            \int_{S_{\min}}^{S_{\max}} 
                V(S) \Big| V(S) - \hat{V}(S;\boldsymbol{\theta}) \Big| dS + 
            \gamma \|\boldsymbol{\theta}\|_p^p
    \end{equation}
    This formulation provides error weighting proportional to the target payout magnitude, 
    offering economic interpretation well-aligned with trading objectives. Furthermore, it 
    naturally enhances focus on regions with significant payout values, making it particularly 
    suitable for practical applications.
\end{enumerate}

However, these generalizations lose the analytical tractability of the L2 case, 
requiring numerical optimization methods for solution.

\subsection{Numerical Implementation}
The implementation leverages the analytical solution derived in the Mathematical Framework 
section for L2-regularized problems. For L1 regularization and alternative loss functions, 
numerical optimization methods are employed.

\begin{figure}[htbp]
\centering
\input{regularization_comparison.tex}
\caption{Comparison of L1 and L2 regularization}
\label{fig:regularization}
\end{figure}

Figure~\ref{fig:regularization} demonstrates the effectiveness of both regularization 
approaches in approximating a complex payout profile. The target function combines 
different types of functions, presenting a challenging test case. Both 
methods achieve good approximation quality, with L2 regularization providing slightly 
smoother results and L1 regularization favoring a sparser representation with fewer 
non-zero option weights.

\section{Results}
\subsection{Distribution-Based Trading}
The approximation method enables novel trading strategies that exploit distributional 
forecasts rather than direct price predictions. Traders can hedge against forecasted 
distribution features, express views on higher moments of price distribution, and 
implement volatility surface arbitrage strategies through carefully constructed 
option portfolios.

The key advantage emerges when combining the approximation with distribution 
forecasts $\mathbb{P}(S_T)$:

\begin{equation}
\mathbb{E}[V(S_T)] = 
    \int_{0}^{\infty} 
        V(S) \mathbb{P}(S_T = S) dS
\end{equation}

Optimal portfolio weights $\boldsymbol{\theta}^*$ maximize the expected payout under 
forecasted distribution:

\begin{equation}
\boldsymbol{\theta}^* = 
    \argmax_{\boldsymbol{\theta}} 
        \mathbb{E}\left[\hat{V}(S_T;\boldsymbol{\theta})\right] - 
        \lambda \text{Var}\left(\hat{V}(S_T;\boldsymbol{\theta})\right)
\end{equation}

where $\lambda$ represents risk aversion parameter.

\section{Conclusion}

\bibliographystyle{apalike}
\bibliography{references}

\end{document}