### 2025-02-19 latex/weighted_loss.tex
weighted

\begin{tikzpicture}
\begin{axis}[
    width=0.9\textwidth,
    height=6cm,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    xlabel style={font=\tiny},
    ylabel style={font=\tiny},
    tick label style={font=\tiny},
    title style={font=\small},
    legend style={font=\tiny, at={(0.02,0.98)}, anchor=north west},
    xlabel={Underlying Asset Price at Maturity},
    ylabel={Payoff},
    title={Weighted Error Method Comparison}
]

\addplot[thick, black] table[x index=0,y index=1] {weighted_loss.dat};
\addlegendentry{Target Payoff}

\addplot[thick, dotted, green!50!black] table[x index=0,y index=2] {weighted_loss.dat};
\addlegendentry{Weighted Error ($\gamma=0.10$)}

\end{axis}
\end{tikzpicture}